% -*- mode: context; -*-
\chapter{Miscellaneous Problems}
\startitemize[n, 1*broad]
\item Two H.M.'s were inserted between two numbers and one of them is equal to the geometric mean between
  the numbers. Prove that one number is equal to four times the other number.
\item The H.M. of two numbers is $4$. The A.M. \quote{A} and G.M. \quote{G} satisfy the relationship $2A +
  G^2 = 27$. Find the numbers.
\item If $n$ is odd find the sum of the series $a - (a + d) + (a + 2d) - (a + 3d) + \cdots$.
\item For what value of $y$, the equation $\cos x = y$ has roots in A.P.
\item Sum of all terms of a G.P.\ is $5$ times the sum of odd terms. Find the common ratio.
\item If the sum of $p$ terms of an A.P.\ be $ap + bp^2$, then find its fourth term.
\item Among $302, 303, 304, 305$, which is a term of the sequence $3, 8, 13, \ldots$?
\item Consider points $(2, 1), (1, -2), (3, 2), (1, 1)$. Now if $A, B, C$ are in A.P. then the straight line
  $Ax + By + C = 0$ will pass through which of the points?
\item If the sum of $n$ terms of an A.P.\ is $\frac{3n^2 + 5n}{36}$, then find the common difference.
\item If $a, b, c$ are three consecutive terms of a sequence such that $\frac{a - b}{b - c} = \frac{a}{c}$,
  then find the type of sequence.
\item Find the values of $x$, for which the sum of the series $\left(\frac{2x}{x + 3}\right)
  + \left(\frac{2x}{x + 3}\right)^2 + \left(\frac{2x}{x + 3}\right)^3 + \cdots$ will be definite.
\item If the product of $n$ positive numbers is $1$, then prove that their sum is never less than $1$.
\item Let $S_n$ denote the sum of first $n$ terms of an A.P. If $S_{2n} = 3S_n$ then find the ratio of
  $S_{3n}/S_n$.
\item Find the sum of $n$ terms of the series $1.3 + 3.5 + 5.7 + \cdots$.
\item Sum the series $1 + \frac{4}{5} + \frac{7}{5^2} + \frac{10}{5^3} + \cdots$ for initite terms.
\item If $a, b, c$ are in A.P.; $a, x, b$ are in G.P.; and $b, y, c$ are in G.P., then find the type of
  sequence $x^2, b^2, y^2$ will form.
\item If $\log_{10}2, \log_{10}(2^x - 1), \log_{10}(2^x + 3)$ are the terms of an A.P., find the value of
  $x$.
\item If the numbers $p, q, r$ are in A.P., then prove that $m^{7p}, m^{7q}, m^{7r}(m > 0)$ are in G.P.x
\item For $0< \phi < \frac{\pi}{2}$, if $x = \displaystyle\sum_{n = 0}^\infty \cos^{2n}\phi, y = \sum_{n =
  0}^\infty\sin^{2n}\phi$ and $z = \displaystyle\sum_{n = 0}^\infty\cos^{2n}\phi\sin^{2n}\phi$, then prove
  that $xyz = xy + z$.
\item Find the sum of the series $\sqrt{2} + \sqrt{8} + \sqrt{18} + \sqrt{32} + \cdots$.
\item In a triangle if lengths of sides $a, b, c$ are in A.P. then prove that
  $\tan\frac{A}{2}, \tan\frac{B}{2}, \tan\frac{C}{2}$ are in H.P.
\item The ex-radii of a triangle are in H.P. Prove that $a, b, c$ are in A.P.
\item If the first and $(2n - 1)$st terms of an A.P., G.P.\ and H.P.\ are equal and their $n$th terms are
  $a, b$ and $c$ then prove that $ac - b^2 = 0$.
\item If the roots of the equation $x^3 - 12x^2 + 39 - 28 = 0$ are in A.P., then find their common
  difference.
\item If $a, b, c, d$ and $p$ are distinct real numbers such that $(a^2 + b^2 + c^2)p^2 - 2(ab + bc + cd)p +
  (b^2 + c^2 + d^2)\leq 0$ then prove that $a, b, c, d$ are in G.P.
\item Four numbers are such that first three are in A.P., while the last three are in G.P. If the first
  number is $6$ and common ratio is of the G.P.\ is $\frac{1}{2}$, then find the numbers.
\item If $\frac{44}{9} = 3 + 5r + 7r^2 + 9r^3 + \cdots$ to $\infty$, then find the value of $r$.
\item If $|x| < 1$, then find the sum of the series $1^2 + 2^2x + 3^2x^2 + 4^2x^3 + \cdots$ to $\infty$.
\item Find the sum of $1^2 + (1^2 + 2^2) + (1^2 + 2^2 + 3^2) + \cdots$ up tp $22$ terms.
\item If the ratio of sum of $m$ and $n$ terms of an A.P.\ is $m^2:n^2$, then find the ratio of its $m$th
  and $n$th term.
\item The sum of $n$ terms of two A.P.\ are in the ratio of $7n + 1:4n + 27$. Find the ratio of their $11$th
  terms.
\item A car travels from station $A$ to station $B$ with a speed of $40$ kmph and returns with a speed of
  $60$ kmph. Find the average speed in kmph.
\item If $a, b, c$ are in A.P.\ and $a^2, b^2, c^2$ are in H.P., then prove that $a = b = c$.
\item Find the sum of series $1 + \frac{1}{1 + 2} + \frac{1}{1 + 2 + 3} + \cdots$ upto $\infty$.
\item If the sum of roots of the quadratic equation $ax^2 + bx + c = 0$ is equal to the sum of the squares
  of the reciprocals, then prove that $ab^2, ca^2, bc^2$ are in A.P.
\item If $p$ times the $p$th term of an A.P.\ is equal to $q$ times the $q$th term of the same A.P., then
  prove that $(p + q)$th terms is zero.
\item If in a G.P.\ the first term is $a$, the second term is $b$ and the last term is $c$, then find the
  sum of the series.
\item In a triangle, the length of the two larger sides are $10$ and $9$ respectively. If the angles are in
  A.P., then find the length of the third side.
\item If one geometric mean $G$ and two arithmetic means $p$ and $q$ are inserted between two quantities,
  then prove that $G^2 = (2p - q)(2q - p)$.
\item If $a, b, c$ are three unequal numbers such that $a, b, c$ are in A.P., and $b - a, c - b, a$ are in
  G.P. then prove that $a:b:c = 1:2:3$.
\item If the first and last terms of an A.P.\ are $a$ and $l$ respectively, and $s$ be the sum of all terms
  in the A.P., then find the common difference.
\item Let $T_r$ be the $r$th term of an A.P. for $r = 1, 2, 3, \ldots$ If for some positive integers $m, n$
  we have $T_m = \frac{1}{n}$ and $T_n = \frac{1}{m}$, then find $T_{mn}$.
\item If $\sin\theta, \cos\theta$ and $\tan\theta$ are in G.P., then find the value of $\cot^6\theta
  - \cot^2\theta$.
\item If the perimeter of a $\triangle ABC$ is $6$ times the arithmetic mean of the sines of its angle and
  the side $a$ is $1$, then find the $\angle A$.
\item Find the sum $1.1! + 2.2! + 3.3! + \cdots + n.n!$.
\item If $x, y, z$ are in A.P., then prove that $\frac{1}{\sqrt{x} + \sqrt{y}}, \frac{1}{\sqrt{z}
  + \sqrt{x}}, \frac{1}{\sqrt{y} + \sqrt{z}}$ are also in A.P.
\item Find the sum of the series $2 + \frac{5}{2!3} + \frac{5.7}{3!3^2} + \frac{5.7.9}{4!3^3} + \cdots$.
\item If $1 + \cos\alpha + \cos^2\alpha + \cdots = 2 - \sqrt{2}$, then find $\alpha(0 < \alpha < \pi)$.
\item Find the H.M.\ of the roots of the equation $(5 + \sqrt{2})x^2 - (4 + \sqrt{5})x + 8 + 2\sqrt{5} = 0$.
\item If $\startdeterminant\NC a\NC b\NC a + b\NR\NC b\NC c\NC b + c\NR\NC a + b\NC b + c\NC
  0\NR\stopdeterminant = 0$, then prove that $a, b, c$ are in G.P.
\item For any three positive real numbers $a, b$ and $c$ if $9(25a^2 + b^2) + 25(c^2 - 3ac) = 15b(3a + c)$,
  then prove that $b, c, a$ are in A.P.
\item Let $AP(a, d)$ denote the set of all terms of an A.P.\ with first term $a$ and common difference $d >
  0$. If $AP(1, 3)\cap AP(2, 5)\cap AP(3, 7) = AP(a, d)$ then find $a$ and $d$.
\item The sides of a right angled triangle are in A.P. If the area of triangle is $24$ then find the length
  of the smallest side.
\item For $x\in\mathbb{R}$, let $[x]$ denote the greatest integer $\leq x$, then find the sum of the series
  $\left[-\frac{1}{3}\right] + \left[-\frac{1}{3} - \frac{1}{100}\right] + \left[-\frac{1}{3}
  - \frac{2}{100}\right] + \cdots + \left[-\frac{1}{3} - \frac{99}{100}\right]$.
\item Let the sun of first $n$ terms of a non-constant A.P.\ be $50n + \frac{n(n - 7)}{2}A$, where $A$ is a
  constant. Find the first term and common difference of this A.P.
\item Find the sum of all two digit positive numbers which when divided by $7$ leaves a remainder of $2$ or
  $5$.
\item Let $a_1, a_2, \ldots, a_{30}$ be in an A.P., $S = \displaystyle\sum_{i = 1}^{30}a_i$ and $T
  = \displaystyle\sum_{i = 1}^{15}a_{2i - 1}$. If $a_5 = 27$ and $S - 2T = 75$, then find $a_{10}$.
\item Let $b_i > 1$ for $i = 1, 2, \ldots, 101$. Suppose $\log_eb_1, \log_eb_2, \ldots, \log_eb_{101}$ are
  in A.P. with the common difference of $\log_e2$. Suppose $a_1, a_2, \ldots, a_{101}$ are in A.P. such that
  $a_1 = b_2$ and $a_{51} = b_{51}$. If $t = b_1 + b_2 + \cdots + b_{51}$ and $s = a_1 + a_2 + \cdots +
  a_{51}$, then prove that $s > t$ and $a_{101} < b_{101}$.
\item If the sum of first $n$ terms of an A.P.\ is $cn^2$, then find the sum of squares of these $n$ terms.
\item Let $V_r$ denote the sum of first $r$ terms of an A.P.\ whose first term is $r$ and the common
  difference is $2r - 1$. Given $T_r = V_{r + 1} - V_r$ and $Q_r = T_{r + 1} - T_r$. Find $V_1 + V_2
  + \cdots + V_n$. Prove that $T_r$ is always a composite number and $Q_1, Q_2, Q_3, \ldots$ form an
  A.P.\ with common difference of $6$.
\item Let $p$ and $q$ be the roots of the equation $x^2 - 2x + A = 0$, and, $r$ and $s$ be the roots of the
  equation $x^2 - 18x + B = 0$. If $p < q < r < s$ form an A.P., then find $A$ and $B$.
\item The sum of the series $1^2 + 2.2^2 + 3^2 + 2.4^2 + 5^2 + 2.6^2 + \cdots$ is $\frac{n(n + 1)^2}{2}$
  when $n$ is even. Find the sum when $n$ is odd.
\item The fourth power of the common difference of an A.P.\ with integer entries is added to the produt of
  four consecutive terms of it. Prove that the resulting sum is the square of an integer.
\item The real numbers $x_1, x_2, x_3$ satisfying the equation $x^3 - x^2 + \beta x + \gamma = 0$ are in
  A.P. Find the intervals in which $\beta$ and $\gamma$ lie.
\item Suppose that all the terms of an A.P.\ are natural numbers. If the ratio of the sum of the first seven
  terms to the sum of the first eleven terms is $6:11$ and the seventh term lies between $130$ and $140$,
  then find the c.d.\ of the A.P.
\item A pack contains $n$ cards numbered from $1$ to $n$. Two consecutive cards are removed from the pack
  and the sum of the remaining cards is $1224$. If the smaller of the numbers on the removed cards is $k$,
  then find $k$.
\item Let $a_1, a_2, a_3, \ldots, a_{100}$ be an A.P.\ with $a_1 = 3$ and $S_p = \displaystyle\sum_{i =
  1}^pa_i, 1\leq p\leq 100$. For any integer $n$ with $1\leq n\leq 20$, let $m = 5n$. If $\frac{S_m}{S_n}$
  does not depend on $n$ then find common difference.
\item Let $a_1, a_2, a_3, \ldots, a_{11}$ be real numbers satisfying $a_1 = 15, 27 - a_2 > 0$ and $a_k =
  2a_{k - 1} - a_{k - 2}$ for $k = 3, 4, \ldots, 11$. If $\frac{a_1^2 + a_2^2 + \cdots + a_{11}^2}{11} =
  90$, then find the value of $\frac{a_1 + a_2 + \cdots + a_{11}}{11}$.
\item Let $a, b$ and $c$ be in G.P. with common ratio $r$, where $a\neq 0$ and $0 < r \leq \frac{1}{2}$. If
  $3a, 7b$ and $15c$ are in A.P., then find the $4$th term of this A.P.
\item If $a, b$ and $c$ are three, distinct real numbers in a G.P.\ such that $a + b + c = xb$, then find
  the value of $x$.
\item Let $f(x) = ax^2 + bx + c, a\neq 0$ and $\Delta = b^2 - 4ac$. If $\alpha + \beta, \alpha^2 + \beta^2$,
  and $\alpha^3 + \beta^3$ are in G.P., ($\alpha, \beta$ are roots of the given quadratic equation) then
  prove that $\Delta = 0$.
\item If $|x| < 1, |y| < 1$ and $x\neq y$, then sum the following series to infinity: $(x + y) + (x^2 + xy +
  y^2) + (x^3 + x^2y + xy^2 + y^2) + \cdots$.
\item Let $S_n = 1 + q + q^2 + \cdots + q^n$ and $T_n = 1 + \left(\frac{q + 1}{2}\right) + \left(\frac{q +
  1}{2}\right)^2 + \cdots + \left(\frac{q + 1}{2}\right)^n$, where $q$ is a real number and $q\neq 1$. If
  $C_1^^{101} + C_2^^{101}S_1 + C_3^^{101}S_2 + \cdots + C_{101}^^{101}S_{100} = \alpha.T_{100}$, then find
  $\alpha$.
\item If $10^9 + 2.11.10^8 + 3.11^2.10^7 + \cdots + 10.11^8 = k.10^9$, then find $k$.
\item Let $\alpha$ and $\beta$ be the roots of $x^2 - x - 1 = 0$, with $\alpha > \beta$. For all positive
  integers $n$ define $a_n = \frac{\alpha^n - \beta^n}{\alpha - \beta}, n\geq 1, b_n = 1$ and $b_n = a_{n -
    1} + a_{n - 2}, n\geq 2$. Prove that
  \startitemize[a]
  \item $b_n = \alpha^n + \beta^n$ for all $n \geq 1$,
  \item $a_1 + a_2 + \cdots + a_n = a_{n + 2} - 1$ for all $n\geq 1$, and
  \item $\displaystyle\sum_{n = 1}^\infty\frac{a_n}{10^n} = \frac{10}{89}$.
  \stopitemize
\item Let $A_n = \frac{3}{4} - \left(\frac{3}{4}\right)^2 + \left(\frac{3}{4}\right)^3 - \cdots + (-1)^{n -
  1}\left(\frac{3}{4}\right)^n, B_n = 1 - A_n$. Find the least odd number $n_0$ such that $B_n >
  A_n, \forall n \geq n_0$.
\item Let $S_k$, where $k = 1, 2, \ldots, 100$, denotes the sum of the infinite geometric series whose first
  terrm is $\frac{k - 1}{k!}$ and common ratio is $\frac{1}{k}$. Find the value of $\frac{100^2}{100!}
  + \displaystyle\sum_{k = 1}^{100}|(k^2 - 3k + 1)S_k|$.
\item If $\cos(x - y), \cos x$ and $\cos(x + y)$ are in H.P., then find the value of $\cos
  x\sec\frac{y}{2}$.
\item Find the sum of the series $1 + \frac{1^3 + 2^3}{1 + 2} + \frac{1^3 + 2^3 + 3^3}{1 + 2 + 3} + \cdots +
  \frac{1^3 + 2^3 + \cdots + 15^3}{1 + 2 + \cdots + 15} - \frac{1}{2}(1 + 2 + \cdots + 15)$.
\item Find the sum of the series $\frac{3\times1^3}{1^2} + \frac{5\times(1^3 + 2^3)}{1^2 + 2^2}
  + \frac{7\times(1^3 + 2^3 + 3^3)}{1^2 + 2^2 + 3^2} + \cdots$ for $10$ terms.
\item If the sum of the series $\left(\frac{3}{4}\right)^3 + \left(1\frac{1}{2}\right)^3
  + \left(2\frac{1}{4}\right)^3 + 3^3 + \left(3\frac{3}{4}\right)^3 + \cdots$ for $15$ terms is $225k$, then
  find $k$.
\item Let $S_k = \frac{1 + 2 + \cdots + k}{k}$. If $S_1^2 + S_2^2 + \cdots + S_{10}^2 = \frac{5}{12}A$, then
  find $A$.
\item Let $x, y$ be positive real numbers and $m, n$ positive integers. Then find the maximum value of the
  expression $\frac{x^my^n}{\left(1 + x^{2m}\right)\left(1 + y^{2n}\right)}$.
\item Find the sum of the series $1 + 6 + \frac{9(1^2 + 2^2 + 3^2)}{7} + \frac{12(1^2 + 2^2 + 3^2 + 4^2)}{9}
  + \frac{15(1^2 + 2^2 + 3^2 + 4^2 + 5^2)}{11} + \cdots$ for $15$ terms.
\item Let $a_1, a_2, \ldots, a_{49}$ be in A.P.\ such that $\displaystyle\sum_{k = 0}^{12}a_{4k + 1} = 416$
  and $a_9 + a_{43} = 66$. If $a_1^2 + a_2^2 + \cdots + a_{17}^2 = 140m$, then find the value of $m$.
\item Let $A$ be the sum of first $20$ terms and $B$ be the sum of $40$ terms of the series $1^2 + 2.2^2 +
  3^2 + 2.4^2 + 5^2 + 2.6^2 + \cdots$. If $B - 2A = 100\lambda$, then find $\lambda$.
\item If the sum of first $10$ terms of the series $\left(1\frac{3}{5}\right)^2
  + \left(2\frac{2}{5}\right)^2 + \left(3\frac{1}{5}\right)^2 + 4^2 + \left(4\frac{4}{5}\right)^2 + \cdots$
  is $\frac{16m}{5}$, then find $m$.
\item If $m$ is the A.M.\ of two distinct real numbers $l$ and $n(l, n > 1)$ and $G_1, G_2$ and $G_3$ are
  three G.M.\ between $l$ and $n$, then find the value of $G_1^4 + 2G_2^4 + G_3^4$ in terms of $l, m$ and
  $n$.
\item Find the sum of first $9$ terms of the series $\frac{1^3}{1} + \frac{1^3 + 2^3}{1 + 3} + \frac{1^3 +
  2^3 + 3^3}{1 + 3 + 5} + \cdots$.
\item For $\alpha\in\left(0, \frac{\pi}{2}\right)$, prove that $\sqrt{x^2 + x}
  + \frac{\tan^2\alpha}{\sqrt{x^2 + x}}$ is always greater than or equal to $2\tan\alpha$.
\item If $a_1, a_2, \ldots, a_n$ are positive real numbers whose product is a fixed number $c$, then find
  the minimum value of $a_1 + a_2 + \cdots + a_{n - 1} + a_{2n}$.
\item If $a, b, c$ are positive real numbers such that $a + b + c + d = 2$, then prove that the relation $M
  = (a + b)(c + d)$ satisfies the relation $0 < M \leq 1$.
\item Find the harmonic mean of the roots of the equation $\left(5 + \sqrt{2}\right)x^2 - \left(4 +
  \sqrt{5}\right)x + 8 + 2\sqrt{5} = 0$.
\item If $a, b, c$ are distinct positive numbers, then prove that the expression $(b + c - a)(c + a - b)(a +
  b - c) - abc$ is negative.
\item If $a, b, c$ are positive real numbers, then prove that $\left[(1 + a)(1 + b)(1 + c)\right]^7 >
  7^7a^4b^4c^4$.
\item Let $a_1, a_2, \ldots$ be positive real numbers in G.P. For each $n$, if $A_n, G_n, H_n$ are
  respectively the A.M., G.M., H.M.\ of $a_1, a_2, \ldots, a_n$. Then find an expression for the G.M.\ of
  $G_1, G_2, \ldots, G_n$ in terms of $A_1, A_2, \ldots, A_n, H_1, H_2, \ldots, H_n$.
\item Let $m$ be the minimum possible value of $\log_3\left(3^{y_1} + 3^{y_2} + 3^{y_3}\right)$, where $y_1,
  y_2, y_3$ are real numbers such that $y_1 + y_2 + y_3 = 9$. Let $M$ be the maximum possible value of
  $\log_3x_1 + \log_3x_2 + \log_3x_3$, where $x_1, x_2, x_3$ are real numbers such that $x_1 + x_2 + x_3 =
  9$. Find the value of $\log_2m^3 + \log_3M^2$.
\item Let $a_1, a_2, a_3, \ldots$ be a sequence of positive integers in A.P.\ with c.d.\ $2$. Also, let
  $b_1, b_2, b_3, \ldots$ be a sequence of positive integers in G.P.\ with c.r.\ $2$. If $a_1 = b_2 = c$,
  then find the number of all possible values of $c$ for which the equality $2(a_1 + a_2 + \cdots + a_n) =
  b_1 + b_2 + \cdots + b_n$ holds.
\item Let $a, b, c$ be positive integers such that $b/a$ is an integer. If $a, b, c$ are in G.P.\ and the
  A.M.\ of $a, b, c$ is $b + 2$, then find the value of $\frac{a^2 + a - 14}{a + 1}$.
\item Find the minimum value of the sum of real numbers $a^{-5}, a^{-4}, 3a^{-3}, 1, a^8$ and $a^{10}$ where
  $a > 0$.
\item Let $a_1 = b_1 = 1; a_n = a_{n - 1} + 2$ and $b_n = a_n + b_{n - 1}$ for every natural number $n\geq
  2$. Find $\displaystyle\sum_{n = 1}^{15}a_n.b_n$.
\item If $\frac{1}{2.3^{10}} + \frac{1}{2^2.3^9} + \cdots + \frac{1}{2^{10}.3} = \frac{K}{2^{10}.3^{10}}$,
  then find the remainder when $K$ is divided by $6$.
\item Find the greatest integer less than or equal to the sum of first $100$ terms of the sequence
  $\frac{1}{3}, \frac{5}{9}, \frac{19}{27}, \frac{65}{81}, \cdots$.
\item Consider two G.P.'s $2, 2^2, 2^3, \ldots$ and $4, 4^2, 4^3, \ldots$ of $60$ and $n$ terms
  respectively. If the G.M.\ of all $60 + n$ terms is $2^{\frac{225}{8}}$, then find $\displaystyle\sum_{k =
    1}^{n}k(n - k)$.
\item If $\displaystyle\sum_{k = 1}^{10}\frac{k}{k^4 + k^2 + 1} = \frac{m}{n}$, where $m$ and $n$ are
  co-prime. Find $m + n$.
\item Different A.P.'s are constructed with the first term $100$ and the last term $199$ with integral
  common difference. Find the sum of common differences of all such A.P.'s having at least $3$ terms and at
  most $33$ terms.
\item If $A = \displaystyle\sum_{n = 1}^\infty\frac{1}{\left(3 + (-1)^n\right)^n}$ and $B =
  \displaystyle\sum_{n = 1}^\infty\frac{(-1)^n}{\left(3 + (-1)^n\right)^n}$, then find $\frac{A}{B}$.
\item If $a_1(> 0), a_2, a_3, a_4, a_5$ are in a G.P., $a_2 + a_4 = 2a_3 + 1$ and $3a_2 + a_3 = 2a_4$, then
  find $a_2 + a_4 + 2a_5$.
\item If the sum of the first ten terms of the series $\frac{1}{5} + \frac{2}{65} + \frac{3}{325} +
  \frac{4}{1025} + \frac{5}{2501} + \cdots$ is $\frac{m}{n}$, where $m, n$ are co-prime, then find $m + n$.
\item Let $S = 2 + \frac{6}{7} + \frac{12}{7^2} + \frac{20}{7^3} + \frac{30}{7^4} + \cdots \infty$, then
  find $4S$.
\item Consider the sequence $a_1, a_2, a_3, \ldots$ such that $a_1 = 1, a_2 = 2$ and $a_{n  + 2} =
  \frac{2}{a_{n + 1}} + a_n$ for $n = 1, 2, 3, \ldots$. If $\left(\frac{a_1 +
    \frac{1}{a_2}}{a_3}\right)\left(\frac{a_2 + \frac{1}{a_3}}{a_4}\right)\left(\frac{a_3 +
    \frac{1}{a_4}}{a_5}\right) \ldots \left(\frac{a_{30} + \frac{1}{a_{31}}}{a_{32}}\right) = 2^\alpha
  C_{31}^^{61}$, then find $\alpha$.
\item Let the coefficients of the middle terms in the expansion of $\left(\frac{1}{\sqrt{6}} + \beta
  x\right)^4, (1 - 3\beta x)^2$ and $\left(1 - \frac{\beta}{2}x\right)^6, \beta > 0$, respectively form the
  first three terms of an A.P. If $d$ is the c.d.\ of this A.P.\ then find the value of $50 -
  \frac{2d}{\beta^2}$.
\item Let $A_1, A_2, A_3, \ldots$ be an increasing G.P.\ of positive real numbers. If $A_1A_3A_5A_7 =
  \frac{1}{1296}$ and $A_2 + A_5 = \frac{7}{36}$, then find the value of $A_6 + A_8 + A_{10}$.
\item If $n$ A.M.\ are inserted between $a$ and $100$ such that the ratio of first mean to the last mean is
  $7$ and $a + n = 33$, then find the value of $n$.
\item Let for $n = 1, 2, 3, \ldots, 50, S_n$ be the sum of the infinite G.P. whose first term is $n^2$ whose
  common ratio is $\frac{1}{(n + 1)^2}$. Find the value of $\frac{1}{26} + \displaystyle\sum_{n =
    1}^{50}\left(S_n + \frac{2}{n + 1} - n - 1\right)$.
\item Let $a_1, a_2, a_3, \ldots$ be an A.P. If $\displaystyle\sum_{r = 1}^\infty\frac{a_r}{2^r} = 4$, then
  find $a_2$.
\stopitemize
