% -*- mode: context; -*-
\chapter{Permutations and Combinations}
\startitemize[n, 1*broad]
\item Given, $P_4^^n = 360 \Rightarrow n(n - 1)(n - 2)(n - 3) = 3\times 4\times 5\times 6\Rightarrow n = 6$.
\item Given, $P_3^^n = 9240 \Rightarrow n(n - 1)(n - 2) = 20\times21\times22 \Rightarrow n = 22$.
\item Given, $P_r^^{10} = 720 = 8\times9\times10 \Rightarrow r = 3$.
\item Given, $P_{n - 1}^^{2n + 1}:P_n^^{2n - 1} = 3: 5 \Rightarrow \frac{(2n + 1)!}{(n + 2)!}.\frac{(2n -
  1)!}{(n -- 1)!} = \frac{3}{5}\Rightarrow \frac{(2n + 1)2n}{n(n + 1)(n + 2)}= \frac{3}{5}$

  $\Rightarrow 3n^2 - 11n - 4 = 0 \Rightarrow n = 4, -\frac{1}{3}$, but $n$ is an integer. Hence, $n = 4$.
\item Given, $P_4^^n = 12\times P_2^^n \Rightarrow n(n - 1)(n - 2)(n - 3) = 12\times n(n - 1)\Rightarrow n^2
  - 5n - 6 = 0 \Rightarrow n = 6, -1$.

  But $n > 0 \Rightarrow n = 6$ is the only solution.
\item Given, $P_5^^n = 20\times P_3^^n \Rightarrow (n -3)(n - 4) = 20  \Rightarrow n^2 - 7n - 8 =
  0\Rightarrow n = 8, -1$.

  But $n > 0 \Rightarrow n = 8$ is the only solution.
\item Given, $P_4^^n:P_4^^{n+1} = 3:4\Rightarrow \frac{n!}{(n - 4)!}.\frac{(n - 3)!}{(n + 1)!} =
  \frac{3}{4}$

  $\Rightarrow \frac{(n - 3)}{n + 1} = \frac{3}{4}\Rightarrow 4n - 12 = 3n + 3 \Rightarrow n = 15$.
\item Given $P_r^^{20} = 6840 = 18\times19\times20 \Rightarrow r = 3$.
\item Given, $P_{k + 1}^^{k + 5} = \frac{11(k - 1)}{2}.P_{k}^^{k + 3} \Rightarrow (k + 5)(k + 4)(k + 3)\cdots
  6.5 = \frac{11(k - 1)}{2}.(k + 3)(k + 2)\cdots 5.4$

  $\Rightarrow (k + 5)(k + 4) = 22k - 22 \Rightarrow k^2 - 13k + 42 = 0 \Rightarrow k = 6, 7$.
\item Given, $P_{r + 1}^^{22}:P_{r + 2}^^{20} = 11:52\Rightarrow \frac{22!}{(21 - r)!}.\frac{(18 - r)!}{20!}
  = \frac{11}{54}$

  $\Rightarrow \frac{22.21}{(21 - r)(20 - r)(19 - r)} = \frac{11}{52} \Rightarrow (21 - r)(20 - r)(19 - r) =
  42.52 = 12. 13.14 \Rightarrow r = 7$.
\item Given, $P_2^^{m + n} = 90 \Rightarrow (m + n)(m + n - 1) = 10.9 \Rightarrow m + n = 10$, and

  $P_2^^{m - n} = 30\Rightarrow (m - n)(m - n - 1) = 6.5\Rightarrow m - n = 6 \Rightarrow m = 8, n = 2$.
\item Given, $P_r^^{12} = 11880 \Rightarrow \frac{12!}{(12 - r)!} = 9\times\10\times11\times 12 \Rightarrow
  r = 4$.
\item Given, $P_{r + 6}^^{56}:P_{r + 3}^^{54} = 30800:1 \Rightarrow \frac{56!}{(50 - r)!}.\frac{(51 -
  r)!}{54!} = 30800$

  $\Rightarrow 56\times55\times(51 - r) = 30800 \Rightarrow 51 - r = 10 \Rightarrow r = 41$.
\item $n.P_n^^n = n.n! = (n + 1 - 1).n! = (n + 1)! - n!$. Similarly, $(n - 1).P_{n - 1}^^{n - 1} = n! - (n -
  1)!, \ldots, 2.P_2^^2 = 3! - 2!, 1.P_1^^1 = 2! - 1!$.

  Adding these, we obtain L.H.S. $= (n + 1)! - 1! = P_{n + 1}^^{n + 1} - 1 =$ R.H.S.
\item Given, $C_{30}^^n = C_4^^n \Rightarrow \frac{n!}{30!(n - 30)!} = \frac{n!}{4!(n - 4)!}$

  Equating $n - 30 = 4$ and $n - 4 = 30$, we obtain $n = 34$ from both.
\item Given, $C_{12}^^n = C_8^^n \Rightarrow \frac{n!}{(n - 12)!12!} = \frac{n!}{(n - 8)!8!} \Rightarrow n -
  12 = 8$ and $n - 8 = 12$. Thus, $n = 20$

  $C_{17}^^{20} = \frac{20!}{17!3!} = \frac{20\times19\times18}{3\times2} = 1140$, and
  $C_{20}^^{22} = \frac{22!}{20!2!} = \frac{22\times21}{2} = 231$.
\item Given, $C_r^^{18} = C_{r + 2}^^18 \Rightarrow \frac{18!}{(18 - r)!r!} = \frac{18!}{(r + 2)!(16 - r)!}
  \Rightarrow 18 - r = r + 2 \Rightarrow r = 8$ and $r = 16 - r \Rightarrow r = 8$.

  $C_6^^r = C_6^^8 = \frac{8!}{6!2!} = 28$.
\item Given, $C_{n- 4}^^n = 15 \Rightarrow \frac{n!}{(n - 4)!4!} = 15 \Rightarrow n(n - 1)(n - 2)(n - 3) =
  3\times 4\times 5\times 6\Rightarrow n = 6$.
\item Given, $C_r^^{15}:C_{r - 1}^^15 = 11:5\Rightarrow \frac{15!}{(15 - r)!r!}.\frac{(r - 1)!(16 -
  r!)}{15!} = \frac{11}{5}\Rightarrow \frac{16 - r}{r} = \frac{11}{5}\Rightarrow r = 5$.
\item Given, $P_r^^n = 2520 \Rightarrow \frac{n!}{(n - r)!} = 2520$ and $C_r^^n = 21 \Rightarrow
  \frac{n!}{(n - r)!r!} = 21$

  $\Rightarrow \frac{2520}{r!} = 21 \Rightarrow r! = 120 \Rightarrow r = 5\Rightarrow n(n - 1)(n - 2)(n -
  3)(n - 4) = 2520 = 7\times 6\times 5\times 4\times 3\Rightarrow n = 7$.
\item We know that $C_r^^n = C_{n - r}^^n\Rightarrow C_{13}^^{20} = C_7^^{20}$ and $C_{14}^^{20} =
  C_6^^{20}$.

  $\therefore C_{13}^^{20} + C_{14}^^{20} - C_6^^{20} - C_7^^{20} = 0$.
\item Given, $C_{r - 1}^^n = 36\Rightarrow \frac{n!}{(n - r + 1)(r - 1)!} = 36, C_r^^n = 84 \Rightarrow
  \frac{n!}{(n - r)!r!} = 84$, and $\frac{n!}{(n - r - 1)!(r + 1)!} = 126$.

  Dividing first two, $\frac{r}{n - r + 1} = \frac{3}{7}\Rightarrow 3n = 10r - 3$, and dividing last two

  $\frac{r + 1}{n - r} = \frac{2}{3}\Rightarrow 2n = 5r + 3$. Solving these two equations, we have $n = 9, r
  = 3$.
\item Thoudand's place can be filled in $5$ ways, hundred's place can be filled in $4$ ways, ten's place can
  be filled in $3$ ways and unit's place can be filled in $2$ ways.

  Thus, total number of $4$ digit numbers is $5\times 4\times 3\times 2 = 120$.

  Alternatively, it is $P_4^^5 = 120$.
\item Hundred's place can be filled in $3$ ways excluding $0, 2, 3$, ten's place can be filled in $5$ ways
  and unit's place can be filled in $4$ ways.

  Thus, no. of numbers between $400$ and $1000$ is $5\times 4\times 3 = 60$.
\item {\bf Case I:} When the number is of three digits i.e. between $300$ and $1000$.

  Hundred's place can be filled in $3$ ways using $3, 4$ or $5$, ten's place can be filled in $5$ ways and
  unit's place can be filled in $4$ ways.

  Thus, total no. of three digit numbers is $5\times 4\times 3 = 60$.

  {\bf Case II:} When the number is of four digits i.e. between $1000$ and $3000$.

  Thousand's place can be filled in $2$ ways using $1$ or $2$. Three remaining places can be filled in
  $P_3^^5$ ways i.e. $60$ ways.

  Therefore, total no. of four digit numbers is $2\times 60 = 120$.

  Thus, total no. of numbers between $300$ and $3000$ is $60 + 120 = 180$.
\item {\bf Case I:} When $2$ is at thousands place.

  Hundred's placec can be filled in $4$ ways using $3, 4, 5, 6$. Two remaining places can be filled in
  $P_2^^5$ i.e. $20$ ways. Number of numbers formed in this case is $4\times 20 = 80$.

  {\bf Case II:} When thousands place is occupied by $3, 4, 5$ or $6$.

  We see that there are four ways to fill thousands place. Three remaining placed can be filled in $P_3^^6$
  i.e. $120$ ways. Number of numbers formed in this case is $4\times 120 = 480$.

  Hence, total no. of numbers is $80 + 480 = 560$.
\item {\bf Case I:} When the number is of one digit.

  There will be four positive numbers excluding $0$.

  {\bf Case II}: When the number is of two digits.

  Ten's place can be filled in $4$ ways using $1, 2, 3$ or $4$. Unit's place can be filled in $P_1^^4$
  ways. Total no. of one digit numbers is $4\times P_1^^4 = 16$.

  {\bf Case III:} When the number is of three digits.

  Hundred's place can be filled in $4$ ways like previous case. Remaining two places can be filled in
  $P_2^^4$ ways. Total no. of three digit numbers is $4\times P_2^^4 = 48$.

  {\bf Case IV:} When the number is four digits.

  Thousand's place can be filled in $4$ ways like previous case. Remaining three places can be filled in
  $P_3^^4$ ways. Total no. of four digit numbers is $4\times P_3^^4 = 96$.

  {\bf Case V:} When the number is of five digits.

  Ten thousand's place can be filled in $4$ ways. Remaining four places can be filled in $P_4^^4$
  ways. Total no. of five digit numbers is $4\times P_4^^4 = 96$.

  Thus, total no. of numbers formed is $4 + 16 + 48 + 96 + 96 = 260$.
\item Total no. of numbers will be $P_4^^4 = 24$. Now since there are $4$ digits and $24$ numbers each
  no. will occur at each place for $6$ times. Thus, sum of digits at each place would be $6(1 + 2 + 3 + 4) =
  60$.

  Therefore, sum of all numbers $60(1 + 10 + 100 + 1000) = 66660$.
\item When any digit except $0$ will occupy unit's place the thousand's place has to be occupied by the
  other two digits. Thus, total no. of such numbers is $3\times2\times P_2^^2 = 12$. Thus, $4$ numbers for
  each of positive digits.

  When one of $1, 2, 3$ occupy thousand's place total no. of numbers is $3\times P_3^^3 = 18$. Thus, $6$
  numbers for each of the positive digits.

  Sum of digits at units, tens and thousands place will be $4(1 + 2 + 3) = 24$ and sum of digits at
  thousands place will be $6(1 + 2 + 3) = 36$.

  Thus, sum of numbers formed is $24(1 + 10 + 100) + 36\times 1000 = 38,664$.
\item Each of the four digits $1, 2, 2, 3$ occurs at each place $\frac{P_3^^3}{2!}$ i.e. $3$ times. Thus,
  sum of digits at each place is $3(1 + 2 + 2 + 3) = 24$.

  Thus, sum of numbers formed $24(1 + 10 + 100 + 1000) = 26,664$.
\item Each friennd can be sent invitation by one servant. Since there are three servants each friend can
  receive an invitaion in three ways. Thus, total no. of ways of sending invitations is $3^6 = 729$.
\item Each prize can be given to any boy. Thus, each prize can be given in $7$ ways, and hence, three
  prizes can be given in $7^3 = 543$ ways.
\item Each arm can occupy four positions, and thus, five arms can have $4^5 = 1024$ ways. But when all arms
  are in rest position no signal can be made. Hence, total no. of signal is $1024 -1 = 1023$ ways.
\item Each ring of lock can have one of the ten letters, then three rings can have $10^3$ combinations of
  the letters. However, one of the combinations will be a successful combination.

  Thus, total no. of possible unsuccessful attempts that can be made is $1000 - 1 = 999$.
\item We have to find numbers which are greater than $1000$ but not greater than $4000$ i.e. $1000< x\leq
  4000$ which is same as $1000\leq x < 4000$.

  Now thousands place can be filled with $1, 2, 3$ i.e. in $3$ ways. Hundreds, tens and units place can be
  filled in $5$ ways each.

  Thus, total no. of numbers which can be formed is $3\times 5^3 = 375$.
\item There are three groups. We can arrange three groups in $3!$ ways. $8$ Indians can be arranged among
  themselves in $8!$ ways, $4$ Ameriacans in $4!$ ways and $4$ Englishmen in $4!$ ways.

  Thus, required answer is $3!8!4!4!$.
\item Total no. of volumes is $4 + 1 + 1 + 1 = 7$. We can arrange these volumes in $7!$ ways. $8$ books
  volume can be arranged in $8!$ ways, volume having $5$ books can be arranged in $5!$ ways and volume of
  $3$ books can be arranged in $3!$ ways.

  Thus, required no. of arrangements is $7!8!5!3!$.
\item Taking all copies of the same book as one, we have $5$ books, which can be arranged in $5!$ ways.

  All copies being identical can be arranged only in $1$ way. Thus, required no. of arrangements is $5! =
  120$.
\item The no. of permutations of the $10$ papers without restriction is $10!$.

  We find our no. of ways in which the best and worst paper come together then subtract from total no. of
  permutations to get the no. of permutations in which they never come together.

  Taking the best and the worst paper as one paper we have $9$ papers, which can be arranged in $9!$ ways,
  but the two papers can be arranged among themselves in $2!$ ways. Thus, total no. of permutatiosn in which
  both the papers are toegther is $9!2!$.

  Thus, no. of permutations in which both are not together is $10! - 9!2! = 8.9!$.
\item Total no. ways in which all of them can be seated is $(5 + 3)! = 8!$. Taking all the girls as one
  total no. of persons is $6$.

  The no. of ways in which these can be seated is $6!$, but the $3$ girls can be arranged in $3!$
  ways. Thus, total no. of ways, when all three girls are together can be seate, is $6!3!$.

  Thus, total no. of ways in which all girls are not together is $8! - 6!3! = 36,000$.
\item Let us first position I.A. students. $*IA*IA*IA*IA*IA*IA*IA*$. The IA indicated the position where
  I.A. students sit and * indicated the positions where I.Sc. students can sit. We observe that there are
  $8$ open places where I.Sc. students can sit.

  Now, $7$ I.A. students can be seated in $7!$ ways and $8$ I.Sc. students can be seated in $P_5^^8$ ways.

  Thus, no. of required arrangements is $7!.\frac{8!}{3!}$.
\item Positioning the boys first, we have $*B*B*B*B*B*B*B*$, where $B$s represents the $7$ boys and $*$s
  represents the open positions for girls.

  $7$ boys can be arranged in $7!$ ways and $3$ girls can be seated in $P_3^^8$ ways. Thus, required no. of
  seating arrangememnts is $7!.\frac{8!}{5!} = 42.8!$.
\item {\bf Case I:} When a boy sits at the first place. The possible arrangement in this case is
  $BGBGBGBG$, where $B$ represents a boy and $G$ represents a girl. Now, $4$ boys and $4$ girls can be
  arranged among themselves in $4!$ ways. Thus, no. of possible seating arrangement in this case is $4!4!$.

  {\bf Case II:} When a girl sits at the first place. Like previous case the possible no. of seating
  arrangements is same i.e. $4!4!$.

  Thus, total no. of seating arrangements is $2.4!4! = 1152$.
\item Possible arrangements will have the form $BGBGBGB$, where $B$ represents a boy, and $G$ represents a
  girl. $4$ boys can be seated in $4!$ ways and $3$ girls can be seated in $3!$ ways.

  Thus, total no. of seating arrangements is $4!3!$.
\item There are $12$ letters in the word civilization; out of which $4$ are i's and other are different.

  Therefore, total no. of permutations is $\frac{12!}{4!}$, which included the word civilization itself.
\item There are $10$ letters in the word university; out of which $4$ are vowels, and {\it i} occurs
  twice. The consonants do not have repetition.

  Treating the $4$ vowels as one letter, because they have to appear together, we have $7$ letters. These
  $7$ letters can be arranged in $7!$ ways. But the four vowels can be arranged among themselves in
  $\frac{4!}{2!}$ ways.

  Thus, total no.of words possible is $7!\frac{4!}{2!}$.
\item There are $8$ letters in the word director; out of which $3$ are vowels, and {\it r} occurs
  twice. Thus, total no. of words is $\frac{8!}{2!}$ .

  When the vowels are together, taking them as one letter, we have $6$ letters, which can be arranged in
  $\frac{6!}{2!}$, but the three vowels can be arranged in $3!$ ways among themselves, making the total
  no. of words in which vowels are together $3!\frac{6!}{2!}$.

  Thus, no. of words in which all three vowels are not together is $\frac{8!}{2!} - 3!\frac{6!}{2!}$.
\item There are $7$ letters in the word welcome; out of which {\it e} occurs twice. Thus, total no. of words
  that can be formed is $\frac{7!}{2!}$.

  If \quote{o} comes at end then we will have $6$ letters left giving us total no. of words as
  $\frac{6!}{2!}$.
\item There are $10$ letters in the word California; out of which $5$ are consonants without repetition and
  $5$ vowels with {\it a} and {\it i} occurring twice.

  Thus, consonants can be arranged in $5!$ ways and vowels can be arranged in $\frac{5!}{2!2!}$ ways.

  Thus, total no. of words possible such that consoanats and vowels occupy their respective places is
  $\frac{5!5!}{2!2!}$.
\item There are $6$ letters in the word pencil with two vowels and three even positions. Thus, vowels can be
  arranged in $P_2^^3 = 6$ ways.

  Rest four positions can be filled in $4! = 24$ ways. Thus, total no. of words is $24\times6 = 144$.
\item From $5$ letters $5!= 120$ words can be formed. Consider the form of word when no two vowels are
  together. $VCVCV$, where {\it C} represents consonants and {\it V} represents the vowels.

  Clearly, consonants can be arranged in $2!$ ways and vowels can be arranged in $P_3^^3 = 3! = 6$ ways.

  Thus, no. of words where vowels are not together is $2\times6 = 12$.
\item There are seven digits given and we have to form numbers greater than one million, which implies all
  seven digits will have to used. Among the given digits $3$ comes thrice and $2$ comes twice. Thus, total
  no. of numbers which can be formed is $\frac{7!}{3!2!} = 420$.

  However, these numbers also contain the numbers where zero is the first digits making them less than one
  million. No. of such numbers is $\frac{6!}{3!2!} = 60$.

  Hence, no. of numbers greater than one million is $420 - 60 = 360$.
\item
  \startitemize[i]
  \item Total no. of persons is $5 + 4 = 9$. With no restirctions they can be seated at a round table in $(9
    - 1)! = 8!$ ways.
  \item Treating all British as a single person because they have to be together we have $6$ persons which
    can be seated in $5!$ ways. But $4$ Britishers can be arranged among themselves in $4!$ ways making the
    total no. of ways $5!4!$.
  \item This is equal to $8! - 5!4!$ from previous parts.
  \item First we seat the $5$ Indians in $4!$ ways. Then that will leave $5$ positions open for Britishers
    between Indians to sit, which gives us $P_4^^5$ ways. Thus, total no. of ways in which no two Britishers
    are together is $4!5!$.
  \stopitemize
\item $5$ Indians can be seated in a circle in $4!$ ways. We will have $5$ positions between Indians in
  which we can seat $5$ Britishers in $P_5^^5 = 5!$ ways.

  Thus, total no. of required ways is $5!4!$.
\item Taking the two delegates who have to always sit together as a single person we have $19$ persons
  which can be seated in $18!$ ways around a round table.

  However, the two delegates themselves can be arranged in $2!$ ways making the required no. of ways
  $18!2!$.
\item No. of four digit numbers which can be formed with $1, 2, 4, 5, 7$ i.e. $5$ digits is $P_4^^5 = 120$.
\item Units place cannot be filled with $0$ so it can be filled in $4$ ways using one of $1, 2, 3, 4$. Rest
  four positions can be filled in $P_4^^4 = 4! = 24$ ways.

  Thus, no. of $5$ digit numbers is $4\times24 = 96$.
\item No. of given digits is $7$ and we have to make numbers between $100$ and $1000$ i.e. three digit
  numbers. Since there is no zero in the given digits the required no. of numbers is $P_3^^7 = 210$.
\item Units place be filled in $5$ ways excluding $0$ and two remaining places can be filled by remaining
  $5$ digits in $P_2^^5 = 20$ ways.

  Thus, total no. of required numbers is $5\times10 = 100$.
\item We have $10$ digits. Units place can be filled in $9$ ways excluding $0$. Rest $8$ places can be
  filled using remaining $9$ digits in $P_8^^9 = 9!$ ways.

  Thus, total no. of $9$ digit numbers with no repetition is $9.9!$.
\item Thousannds place can be filled in $5$ ways excluding $0$. Rest three places can be filled using
  remaining $5$ digits in $P_3^^5 = 60$ ways.

  Thus, no. of required numbers is $5\times60 = 300$.
\item Thousands place can be filled in $2$ ways using either $5$ or $9$. Rest three places can be filled in
  $3!$ ways using remaining three digits.

  Thus, no. of required numbers is $2.3! = 12$.
\item {\bf Case I:} When the number is of three digits.

  Hundreds place can be filled in $3$ ways using $3, 4$ or $5$. Remaining two places can be filled in
  $P_2^^5 = 20$ ways using remaining $5$ digits.

  Thus, no. of three digit numbers is $3\times20 = 60$.

  {\bf Case II:} When the number is of four digits.

  Thousands place can be filled in $3$ ways using $1, 2$ or $3$. Remaining three place can be filled
  in $P_3^^5 = 60$ ways using remaining $5$ digits.

  Thus, no. of four digit numbers is $3\times60 = 180$.

  Thus, no. of required numbers is $60 + 180 = 240$.
\item Since the number has to be divisible by $5$ the units place digit has to be either $0$ or $5$.

  {\bf Case I:} When $0$ is at units place. Rest three places can be filled in $P_3^^4 = 24$ ways using
  remaining $4$ digits.

  Thus, no. of four digit numbers in this case is $24$.

  {\bf Case II:} When $5$ is at units place. Thousands place can be filled in $3$ ways using $4, 6$ or
  $7$. Remaining three places can be filled in $P_2^^3 = 6$ ways using remaining $3$ digits.

  Thus, no. of four digit numbers in this case is $3\times6 = 18$.

  Hence, total no. of required numbers is $18 + 24 = 42$.
\item Since the number has to be even, therefore, units place can be filled by either $2$ or $4$ i.e. in $2$
  ways. Rest four places can be filled in $P_4^^4 = 4! = 24$ ways.

  Thus, total no. of $5$ digit numbers is $2\times24 = 48$.
\item Since the no. has to be divisible by $5$ units place can be occupied only by $0$ and $5$.

  {\bf Case I:} When the no. is of one digit. There are two such numbers $0$ and $5$.

  {\bf Case II:} When the no. is of two digits. If $0$ occurs at units place then tens place can be filled
  in $9$ ways giving us $9$ numbers. However, when $5$ occurs at units place then tens place can be filled
  in $8$ ways giving us $8$ numbers. Thus, total no. of two digits numbers is $17$.

  {\bf Caae III:} When the no. is of three digits. If $0$ occurs at units place then remaining two places
  can be filled in $P_2^^9 = 72$ ways. If $5$ is at units place then hundreds place can be filled in $8$
  ways excluding zero and tens place can be filled in $8$ ways using remaining $8$ digits. Thus, in this
  case otal no. of numbers is $72 + 8\times 8 = 136$.

  Thus, total no. of numbers is $2 + 17 + 136 = 155$.
\item Hundreds place can be filled in $5$ ways excluding $0$. Rest of two places can be filled in $P_2^^5 =
  20$ ways.

  Thus, total no. of numbers is $5\times20 = 100$.

  For odd numbers, units place can be filled in $2$ ways using $5$ or $7$. Hundreds place can be filled in
  $4$ ways excluding $0$ and units place can also be filled in $4$ ways using remaining digits.

  Thus, total no. of odd numbers is $2\times4\times4 = 32$.
\item {\bf Case I:} When the no. is of one digit. There are three such numbers $0, 2$ and $4$.

  {\bf Case II:} When the no. is of two digits. When units place is occupied by $0$, tens place can be filled
  in $4$ ways, making no. of such numbers $4$. If units place is occupied by $2$ or $4$ i.e. in two ways then
  tens place can be filled in $3$ ways excluding $0$, making no. of such numbers $2\times3 = 6$.

  Thus, no. of two digit numbers is $4 + 6 = 10$.

  {\bf Case III:} When the no. is of three digits. When units place is occupied by $0$, remaining two
  places can be filled in $P_2^^4 = 12$ ways, making no. of such numbers $12$. If units place is occupied by
  $2$ or $4$ i.e. in two ways then hundreds place can be filled in $3$ ways excluding $0$ and tens place can
  be filled in $3$ ways using remaining three digits, making no. of such numbers $2\times3\times3 = 18$.

  Thus, no. of three digit numbers is $12 + 18 = 30$.

  {\bf Case IV:} When the no. if of four digits. When units place is occupied by $0$, remaining three places
  can be filled in $P_3^^4 = 24$ ways, making no. of such numbers $24$. Following similarly, when units
  place is occupied by $2$ or $4$, no. of such numbers is $2\times3\times3\times2 = 36$.

  Thus, no. of four digit numbers is $24 + 36 = 60$.

  {\bf Case V:} When the no. is of five digits. In this case, units place must be occupied by $0$ and not by
  $2$ or $4$. Then remaining $4$ places can be filled in $P_4^^4 = 24$ ways.

  Thus, total no. of even numbers is $3 + 10 + 30 + 60 + 24 = 127$.
\item Once we fix $5$ at tens place we have $5$ open places and $5$ different digits, which can be arranged
  in $P_5^^5$ ways.

  Thus, no. of required numbers is $120$.
\item We have $7$ digits, and have to form four digit numbers. No. of such numbers possible is $P_4^^7 =
  840$.

  We have to find numbers greater than $3400$. First we compute numbers between $3400$ and $4000$. The
  thousands place can be filled only by $3$ and hundreds place can be filled by $4, 5, 6$ and $7$ i.e. $4$
  ways. Remaining two positions can be filled in $P_2^^5 = 20$ ways. Thus, no. of numbers between $3400$
  and $4000$ is $4\times20 = 80$.

  Now we compute numbers greater than $4000$. Thousands place can be filled by $4, 5, 6$ and $7$ i.e. in
  $4$ ways. Rest three places can be filled in $P_3^^6 = 120$ ways. Thus, no. of such numbers is $4\times120
  = 480$.

  Thus, no. of numbers greater than $3400$ is $80 + 480 = 560$.
\item Since positions of $3$ and $5$ are fixed rest two positions can be filled with three remaining digits
  in $P_2^^3 = 6$ ways. Thus, no. of such numbers is $6$.
\item Thousands place can be filled in $5$ ways excluding $0$. Rremaining three places can be filled in
  $P_3^^5 = 60$ ways using the five remaining digits. Thus, total no. of four digit numbers is $5\times60 =
  300$.

  For numbers to be greater than $3000$, thousands place has to be filled by $3, 4$ and $5$ i.e. $3$
  ways. Remaining three places can be filled in $P_3^^5 = 60$. Thus, no. of numbers greater than $3000$ is
  $3\times60 = 180$.
\item {\bf Case I:} When the no. is of one digit. Total no. of numbers possible in this case is $7$
  including $0$.

  {\bf Case II:} When the no. is of two digits. Tens place can be filled in $6$ ways excluding $0$ and units
  place can be filled in $6$ ways with remaining digits.

  Thus, no. of two digit numbers is $6\times6 = 36$.

  {\bf Case III:} When no. is of three digits. Following similarly the no. of numbers is $6\times6\times5 =
  180$.

  {\bf Case IV:} When the no. is of four digits. Following similarly the no. of numbers is
  $6\times6\times5\times4 = 720$.

  {\bf Case V:} When the no. is of five digits. Following similarly the no. of numbers is
  $6\times6\times5\times4\times3 = 2160$.

  {\bf Case VI:} When the no. is of six digits. Following similarly the no. of numbers is
  $6\times6\times5\times4\times3\times2 = 4320$.

  {\bf Case VII:} When the no. is of seven digits. Following similarly the no. of numbers is
  $6\times6\times5\times4\times3\times2\times1 = 4320$.

  Thus, total no. of numbers is $7 + 36 + 180 + 720 + 2160 + 4320 + 4320 = 11743$
\item We have $5$ digits so when all of them are taken at a time then no. of possible numbers is $P_5^^5 = 120$.

  Each digit will occupy each place for $24$ numbers. Thus, sum of all numbers at any place is $24(1 + 3 + 5
  + 7 + 9) = 600$. Therefore, sum of all such numbers is $600(1 + 10 + 100 + 1000 + 10000) = 6,666,600$.
\item We have $4$ digits with $3$ occurring twice. Thus, total no. of numbers is $\frac{P_4^^4}{2!} =
  12$. Now each of the digits will occur at each place $\frac{12}{4} = 3$ times.

  Thus, sum of digits at each place is $3(3 + 2 + 3 + 4) = 36$. Thus, sum of all possible numbers is $36(1 +
  10 + 100 + 1000) = 39,996$.
\item Let us fix $2$ at units place. Then, ten thousands place can be filled in $3$ ways using $4, 6, 8$ and
  remaining two places can be filled in $P_3^^3 = 3!$ ways. Thus, total no. of numbers is $3\times6 = 18$.

  Number of numbers when $2$ is at ten throusands place is $P_4^^4 = 24$. Thus, each positive digit will
  occur at units, tens, hundreds and thousands place $18$ times and at thousands place $24$ times.

  Sum of the digits at units, tens, hundreds and thousands place will be each $18(2 + 4 + 6 + 8) = 360$ and
  sum of digits at ten thousands place is $24(2 + 4 + 6 + 8) = 480$.

  Thus, sum of all numbers will be $360(1 + 10 + 100 + 1000) + 480\times10000 = 5,199,960$.
\item Total no. of five digit numbers possible is $P_5^^5 = 120$ where each digit will appear at each
  position $\frac{120}{5} = 24$ times.

  Thus, sum of digits at each place is $24(3 + 4 + 5 + 6 + 7) = 600$. Therefore, sum of all such numbers is
  $600(1 + 10 + 100 + 1000 + 10000) = 6,666,600$.
\item Let us fix $2$ at units place. Then, thousands place can be filled in $2$ ways using $3$ or $5$ and
  remaining two places can be filled in $P_2^^2 = 2$ ways. Thus, total no. of numbers is $2\times 2 = 4$.

  Number of numbers when $2$ is at thousands place is $P_3^^3 = 6$. Thus, each positive digit will
  occur at units, tens, hundreds and thousands place $4$ times and at thousands place $6$ times.

  Sum of digits at units, tens and hundreds place will eb each $4(2 + 3 + 5) = 40$ and sum of digits at
  thousands place will be $6(2 + 3 + 5) = 60$.

  Thus, sum of all numbers will be $40(1 + 10 + 100) + 60\times1000 = 64,440$.
\item Each letter can be put in any one of the four letter boxes. Thus, $5$ letters can be posted in $4^5$
  ways.
\item Each prize can be given in $5$ ways. So three prizes can be given in $5^3$ ways.
\item Each thing can be given in $p$ ways to $p$ person. Thus, $n$ things can be given in $p^n$ ways.
\item Each monkey can have a master in $m$ ways. Thus, $n$ monkeys can have a master in $m^n$ ways.
\item First prize in mathematics and physics can be given in $10$ ways and second prize in $9$ ways. In
  chemistry, first prize can be given in $10$ ways.

  Thus, total no. of ways is $10\times9\times10\times9\times10 = 81,000$.
\item The first animal can be picked in $3$ ways with the possibility of it being a cow, a calf or a
  horse. Similarly, second animal can be picked in $3$ ways. Proceeding this way all $12$ animals for the
  stall can be picked in $3$ ways.

  Thus, total no. of making the shipload is $3^{12}$.
\item Each delegate can be put in a hotel in $6$ ways. Therefore, $5$ delegates can be put in $6^5$ ways.
\item Ten thousands place can be filled in $4$ ways exluding $0$. Rest $4$ places can be filled in $5$ ways
  each. Thus, total no. of $5$ digits numbers is $4\times5^4 = 2,500$.
\item Each ring can be put in a finger in $4$ ways i.e. by putting it in any finger. Thus, $6$ rings can be
  put in $4$ fingers in $4^6$ ways.
\item Thousands place can be filled in $3$ ways using $3, 4$ or $5$. Remaining places can be filled in $6^3$
  ways using any of the digits. But one of these numbers will be $3000$ itself.

  Thus, no. of four digit numbers which can be made is $3\times6^3 - 1$.
\item When the number plate is of three digits, each place can be filled in $9$ ways excluding zero. This
  gives us $9^3$ number plates. Similalry, when the number plate is of four digits the no. of possible
  number plates is $9^4$.

  Thus, total no. of number plates is $9^3 + 9^4 = 10\times9^3 = 7,290$.
\item Each question can be answered in $4$ ways, therefore, $10$ questions can be answered in $4^{10}$ ways.

  Second part: First question can be answered in $4$ ways. Now this choice won't be available for the second
  answer so there are $3$ ways. Similarly, for third and so on. Thus, total no. of ways is $4\times3^9$.
\item Treating all volumes of a book as one book we have four books which can be arranged in $4!$
  ways. However, books having $3$ volumes can be arranged in $3!$ ways among themselves and similarly books
  having $2$ volumes can be arranged in $2!$ ways among themselves.

  Thus, total no. of arranging given books is $4!3!3!2!2!$.
\item There are $14$ books having different no. of copies. Treating all copies as one book we still have
  $14$ books which can be arranged in $14!$ ways.

  Since copies are identical there is only one way to arrange them among themselves. Thus, total no. of
  arranging the given books is $14!$.
\item Treating people of different nationalities as one person we have three persons, which can be arranged
  in $3!$ ways. Now $10$ Indians can be arranged in $10!$ ways among themselves, $5$ Americans can be
  arranged in $5!$ ways among themselves and $5$ Britished can be arranged in $5!$ ways as well.

  Thus, total no. of ways of seating them is $3!10!5!5!$.
\item The pattern would be $GBGBGBGBGBGBG$ where {\it B} shows boys position and {\it G} indicates possible
  positions of girls. Boys can be arranged in $6!$ ways. For girls, there are $7$ open positions and $4$
  girls can be seated in $P_4^^7 = \frac{7!}{3!}$ ways.

  Thus, total no. of ways of seating them is $6!.\frac{7!}{31}$.
\item $n$ books can be arranged in $n!$ ways. Now we will find the no. of arrangements when two given books
  which do not have to be together are together. Treating the two books as one book we have $n - 1$ books
  which can be arraned in $(n - 1)!$ ways. But the two books can be arranged in $2$ ways among themselves,
  making the total no. of arrangements is $2.(n - 1)!$.

  Thus, no. of arrangements when the two books are not together is $n! - 2.(n - 1)! = (n - 2).(n - 1)!$.
\item From previous problem, we find the answer to be $4.5! = 480$.
\item Following like previous problem, we find theh answer to be $480$.
\item Following like previous problem on boys and girls we first seat the $15$ I.Sc.\ students in $15!$ ways
  which gives us $16$ open positions for $B.Sc.$\ students, which can be seated in $P_{12}^^{16}$.

  Thus, total no. of ways of seating the students is $15!.P_{12}^^{16}$.
\item First we arrange black balls which will give us $20$ positions in between them and on the edges for
  white balls. Since the balls are identical we can choose $18$ positions out of $20$ for white balls in
  $C_{18}^^{20} = 190$ ways.
\item First we place $p$ positive signs which will give us $p + 1$ positions for negative signs between them
  and on the edges. Since signs are identical we can choose $n$ positions out of $p + 1$ in $C_n^^{p + 1}$
  ways.
\item $m$ men can be seated in $m!$ ways which will have $m + 1$ positions between them and on the edges for
  women so that no two women sit together. Now $n$ women can be arranged in these $m + 1$ positions in
  $P_n^^{m + 1} = \frac{(m + 1)!}{(m - n + 1)!}$ ways.

  Thus, total no. of ways to seat them is $\frac{m!(m + 1)!}{(m - n + 1)!}$.
\item Following like previous problem, we have $m = 5, n = 3$, so the answer woulld be $\frac{5!6!}{3!}$.
\item We have $12$ alphabets excluding c{\symbol[rightquote]}s out of which $5$ are a{\symbol[rightquote]}s,
  $3$ are b{\symbol[rightquote]}s, $1$ d, $2$ e{\symbol[rightquote]}s and $1$ f,
  so these can be arranged in $\frac{12!}{5!3!2!}$ ways. Now these $12$ alphabets will create $13$ positions
  between them and on the edges which are to be filled by $3$ c's in $P_3^^{13}$ ways.

  Thus, total no. of arrangements is $\frac{12!}{5!3!2!}\times\frac{13!}{10!}$.
\item The word banana has {\symbol[leftquote]}a{\symbol[rightquote]} repeating $3$ times and
  {\symbol[leftquote]}n{\symbol[rightquote]} repeating twice while total no. of alphabets is $6$.

  Hence, to no. of different permutations is $\frac{6!}{3!2!}$.
\item There are $13$ alphabets in the word \quotation{circumference}. \quote{c} comes thrice, \quote{r}
  comes twice, \quote{e} comes thrice and rest come once.

  Thus, total no. of words that can be made is $\frac{13!}{3!3!2!}$.
\item Three copies of four books means $12$ books with repetition of copies. Thus, total no. of arragements
  on the shelf is $\frac{12!}{3!3!3!3!}$.
\item There are $12$ alphabets in the word \quotation{Independence}. \quote{n} comes thrice, \quote{d} comes
  twice, \quote{e} comes four times, and rest come once.

  Thus, total no. of words that can be made is $\frac{12!}{4!3!2!}$.
\item There are $8$ alphabets in the word \quotation{Principal}, of which, \quote{p} comes twice, \quote{i}
  comes twice and rest occur once. Treating all vowels as one alphabet we have $6$ alphabets which can be
  arranged in $\frac{6!}{2!}$ ways.

  However, the vowels themselves can be arranged among themselves in $\frac{3!}{2!}$ ways. Thus, total
  no. of words is $\frac{6!3!}{2!2!}$.
\item There are $11$ alphabets in the word \quotation{Mathematics}, of  which, \quote{m} comes twice,
  \quote{a} comes twice, \quote{t} comes twice and rest comes once. Thus, no. of words that can be formed is
  $\frac{11!}{2!2!2!}$.

  Treating all vowels as one alphabet and all consonants as another we have two alphabets which can be
  arranged in $2!$ ways. But $4$ vowels can be arranged in $\frac{4!}{2!}$ ways and $7$ consonants can be
  arranged in $\frac{7!}{2!2!}$ ways.

  Thus, total no. of such words is $\frac{2!7!4!}{2!2!2!}$.
\item There are $8$ alphabets in the word \quotation{Director}, of which, $r$ comes twice and rest come
  once. Since the vowels have to come together, therefore we treat them as one alphabet making a total of
  $6$ alphabets which can be arranged in $\frac{6!}{2!}$ ways.

  However, the three vowels can be arranged in $3!$ ways among themselves making no. of such words
  $\frac{6!3!}{2!}$.
\item There are $8$ alphabets in the word \quotation{Plantain}, of which, \quote{a} and \quote{n} come
  twice and rest come once. Since the vowels have to come together, therefore we treat them as one alphabet
  making a total of $6$ alphabets which can be arranged in $\frac{6!}{2!}$ ways.

  However, the three vowels can be arraned in $\frac{3!}{2!}$ ways among themselves making no. of such words
  $\frac{6!3!}{2!2!}$.
\item There are $12$ letters in the word \quotation{Intermediate}, of which, \quote{e} comes thrice,
  \quote{i} and \quote{t} comes twice and rest come once.

  We first arrange vowels which can be done in $\frac{6!}{3!2!}$. Now because relative order does not change
  we have six positions for consonants giving us total no. of ways of arranging them as $\frac{6!}{2!}$.

  Thus, total no. of such words is $\frac{6!6!}{3!2!2!}$.
\item There are $8$ letters in the word \quotation{Parallel}, of which, \quote{a} comes twice, \quote{l}
  comes thrice and rest comes once.

  Total no. of arrangements is $\frac{8!}{3!2!}$. Treating all the ls as one letter we have $6$ letters
  which can be arranged in $\frac{6!}{2!}$ ways in which all ls will be together.

  Therefore, no. of words in which all ls are not together is $\frac{8!}{3!2!} - \frac{6!}{2!} = 3000$.
\item The parts are solved below:
  \startitemize[i]
  \item Fixing \quote{D} at the first position; rest four positions can be filled in $P_4^^4$ ways. Thus,
    no. of such words is $4! = 24$.
  \item Fixing \quote{I} at the end; rest four positions can be filled in $P_4^^4$ ways. Thus,
    no. of such words is $4! = 24$.
  \item Fixing \quote{l} in the middle; rest four positions can be filled in $P_4^^4$ ways. Thus,
    no. of such words is $4! = 24$.
  \item Fixing \quote{D} and \quote{I}; rest three positions can be fillled in $P_3^^3$ ways. Thus, no. of
    such words is $3! = 6$.
  \stopitemize
\item There are $7$ unique letter in the word \quotation{Violent} with $3$ vowels. There are $4$ odd places
  so three vowels can be arranged in $P_3^^4 = 4!$ ways. Rest $4$ consonants can be arrannged in $4! = 24$
  ways. Thus, total no. of such words is $24\times 24 = 576$.
\item There are $3$ distinct consonants and $3$ vowels, where \quote{o} repeats once in the word
  \quotation{Saloon}. Since consonants and vowels have to occupy alternate place we will have two
  patterns. $VCVCVC$ and $CVCVCV$, where $C$ represents consonants and $V$ represents vowels.

  Three consonants can be arranged in $3!$ arrangements and $3$ vowels can be arranged in $\frac{3!}{2!}$
  arrangement. Thus, total no. of arrangements is $3!3! = 36$.
\item There are $4$ consonants and $3$ vowels in the word \quotation{Article}. Clearly, there are three even
  places which are to be occupied by vowels in $3!$ arrangements and consonants can be arranged in $4!$
  arrangements for remaining $4$ positions.

  Thus, total no. of words is $4!3! = 144$.
\item Since the number has to be greater than $4$ million and we are given $7$ digits the ten millions place
  can be occupied by either $4$ or $5$ in $2$ ways.

  Remaning digits can be arranged in $\frac{6!}{2!2!} = 180$ arrangements as $2$ and $3$ repeat once. Thus,
  total no. of required numbers is $2\times180 = 360$.
\item In the given digits $2$ comes thrice and $3$ comes twicec so the no. of numbers is $\frac{7!}{3!2!} =
  420$.

  For odd numbers units place is to be occupied by $1, 3$ or $5$. When $1$ or $5$ occupy units place
  remaining positions can be filled in $\frac{6!}{3!2!} = 60$ ways making the number $2\times60 = 120$.

  When one of the $3$'s occupy units place rest of the positions can eb filled in $\frac{6!}{3!} = 120$
  ways. Thus, total no. of odd numbers is $120 + 120 = 240$.
\item There are four odd digits with both $1$ and $3$ repeating. The even no. $2$ repeats once. In a $7$
  digits number there are four odd places which can be filled by odd numbers in $\frac{4!}{2!2!} = 6$ ways.

  Even places can be filled by $2$ and $4$ can be filled in $\frac{3!}{2!} = 3$ ways. Thus, no. of required
  numbers is $6\times3 = 18$.
\item {\bf Case I:} When the no. if is five digits.

  When ten thousands place is occupied by $2, 3$ or $4$ remaining four places can be filled in
  $\frac{P_4^^5}{2!} = 60$ ways, making such numbers $60\times3 = 180$ in number.

  When ten thousands place is occupied by $1$ remaining four places can be filled in $P_4^^5 = 120$ ways.

  Thus, total no. of five digit numbers is $180 + 120 = 300$.

  {\bf Case II:} When the no. is of six digits.

  When hundred thousands place is occupied by $2, 3$ or $4$ remaining five places can be filled in
  $\frac{P_5^^5}{2!} = 60$ ways, making such numbers $60\times3 = 180$ in number.

  When hundred thousands place is occupied by $1$ remaining four places can be filled in $P_5^^5 = 120$ ways.

  Thus, total no. of six digit numbers is $180 + 120 = 300$.

  Thus, total no. of numbers is $300 + 300 = 600$.
\item When the digits are repeated thousands place can be filled in $5$ ways excluding $0$. Remaining $3$
  positions can be filled by $6$ digits in $6^3$ ways.

  Thus, no. of such numbers is $5\times6^3 = 1080$.

  To find the no. of numbers where at least one digit is repeated we find the no. of numbers where no digit
  is repeated and subtract it from previously obtained result.

  For no repetition, thousands placec can be filled in $5$ ways exluding $0$. Remaning $3$ places can be
  filled by $5$ digits in $P_3^^5 = 60$ ways.

  Thus, no. of numbers without repetition is $60\times 5 = 300$.

  Thus, no. of numbers where at least one digit is repeated is $1080 - 300 = 780$.
\item There are a total of $9$ flags, of which, $2$ are red, $2$ are blue and $5$ are yellow. Thus, total
  no. of signals that can be made by using all of them at the same time is $\frac{9!}{2!2!5!}$.
\item When all are of same color $P_1^^6$ signals can be made. When all are of two colors $P_2^^6$ signals
  can be made and so on.

  Thus, total no. of signals is $P_1^^6 + P_2^^6 + P_3^^6 + P_4^^6 + P_5^^6 + P_6^^6 = 1956$.
\item {\bf Case I:} When \quote{e} is in first place. Remaining four places can be filled in $4!$ ways.

  {\bf Case II:} When \quote{e} is in second place. First place can be filled in $3$ ways and remaining $3$
  places in $3!$ ways.

  {\bf Case III:} When \quote{e} is in third place. First two places in $3\times2$ ways and remaining two
  places in $2!$ ways.

  {\bf Case IV:} When \quote{e} is in fourth place. First three places in $3!$ ways and last place with
  \quote{i}.

  Thus, total no. of words is $4! + 3\times3! + 6\times2! + 3! = 60$.

  {\bf Second method:} Total no. of words is $5!$. In half of these \quote{e} will come before \quote{i} and
  in half of them after it. Thus, no. of words is $\frac{5!}{2} = 60$.
\item No. of ways in which $5$ men can sit around a round table is $(5 - 1)! = 24$ arrangements.
\item When there is no restriction we have $10$ girls and boys. Thus, total no. of arrangements would be
  $9!$.

  When no girls are to sit together we first seat the boys in $4!$ arrangements giving us five open
  positions. These can be filled by $5$ girls in $5!$ ways.

  Thus, total no. of seating arrangements is $4!5!$.
\item Treating all girls as a single girl we have $7$ boys and girls which can be seated in $6!$ ways. But
  the $4$ girls can be arranged in $4!$ ways among themselves.

  Thus, total no. of seating arrangements is $6!4!$.
\item The line can start with boys so we first seat the boys put the boys in $5!$ ways followed by girls in
  between boys in $5!$ ways. This can be repeated starting with girls in same manner.

  Thus, no. of lines that can be formed is $2.5!5!$.

  For a round table we have already solved previously giving us $4!5!$ no. of arrangements.
\item $6$ boys can be seated first in $5!$ ways giving us $6$ open places in which girls can be seated in
  $P_5^^6$ ways. Thus, total no. of seating arrangements is $5!6!$.
\item Since in a necklace clockwise and anticlockwise does not matter, therefore, total no. of necklaces
  that can be made using $50$ pearls is $\frac{49!}{2!}$.
\item Treating the two particular delegates as one delegate we have $19$ delegates which can be seated in
  $18!$ ways. But the two delegates can be seated in $2!$ ways among themselves.

  Thus, total no. of seating arrangements is $18!2!$.
\item The question effectively asks for alternate seating arrangements among gentlemen and ladies. Thus,
  followin from problem solved previously total no. of seating arrangements would be $4!3!$.
\item $7$ Englishmen can be seated in $6!$ ways giving us $7$ open places which can be filled by $6$ Indians
  in $P_6^^7$ ways.

  Thus, total no. of seating arrangements is $6!7!$.
\item We know that if $C_x^^n = C_y^^n$ then either $x = y$ or $x + y = n$. Given, $C_{3r}^^{15} = C_{r +
  3}^^{15}$ therefore either $3r = r + 3$ or $3r + r + 3 = 15$.

  However, $3r = r = 3 \Rightarrow r = \frac{3}{2}$, which is not possible, therefore, $3r + r + 3 = 15
  \Rightarrow r = 3$ must be the case.
\item Given, $C_6^^n:C_3^^{n - 3} \Rightarrow \frac{n!}{6!(n - 6!)}.\frac{3!(n - 6)!}{(n - 3)!} =
  \frac{33}{4}$

  $\Rightarrow \frac{n!}{(n - 3)!}.\frac{3!}{6!} = \frac{n(n - 1)(n - 2)}{6.5.4} = \frac{33}{4} \Rightarrow
  n(n - 1)(n - 3) = 11.10.9 \Rightarrow n = 11$.
\item Given, $\displaystyle C_4^^{47} + \sum_{j=1}^5C_3^^{52-j}$

  $= C_4^^{47} + (C_3^^{51} + C_3^^{50} + C_3^^{49} + C_3^^{48} + C_3^^{47}) = (C_4^^{47} + C_3^^{47}) +
  (C_3^^{51} + C_3^^{50} + C_3^^{49} + C_3^^{48})$

  $= C_4^^48 + (C_3^^{51} + C_3^^{50} + C_3^^{49} + C_3^^{48} + C_3^^{47}) [\because C_r^^n + C_{r + 1}^^n =
  C_{r + 1}^^{n + 1}]$

  Repeating this we have the expression equal to $C_4^^{52}$.
\item Let $p$ be the product of $r$ consecutive integers starting from $n$. Then, $p = n(n + 1)(n + 2)\cdots
  (n + r - 1)$

  $\Rightarrow \frac{p}{r!} = \frac{n(n + 1)(n + 2)\cdots(n + r - 1)}{r!} = \frac{1.2.3.\ldots.(n - 1)n(n +
  1)(n + 2)\cdots(n + r - 1)}{1.2.3.\ldots.(n - 1).r!}$

  $= \frac{(n + r - 1)!}{(n - 1)!r!} = C_r^^{n + r - 1}$, which would be an integer, and hence, $p$ is
  divisible by $r!$.
\item A triangle is formed with three vertices so the problem is essentially about choosing $3$ out of $m$
  i.e. $C_3^^m = \frac{m(m - 1)(m - 2)}{6}$.
\item Number of children is $8$. No. of children to be taken at a time is $3$. Out of $8$ children $3$ can
  be selected in $C_3^^8$ ways. Hence, the man has to go to zoo $C_3^^8 = 56$ times.

  Number of selection of $3$ children out of $8$ children including a particular child is $1\times C_2^^7 =
  21$. Hence, a particular child will go $21$ times to the zoo.
\item Let there be $n$ students. No. of ways in which $2$ students can be selected out of $n$ is $C_2^^n$
  i.e. we have $C_2^^n$ pairs.

  But, for each pair of students no. of cards sent is $2$. Thus, total no. of cards sent is $2.C_2^^n = n(n
  - 1) = 600 \Rightarrow n = 25$ because $n\neq -24$.

  {\bf Second method:} Each student sends cards to $n - 1$ students. Thus, total no. of cards sent is $n(n -
  1) = 600 \Rightarrow n = 25$.
\item A polygon of $m$ sides will have $m$ vertices. When any two vertices of the polygon are joined, either
  a diagonal or a side is formed.

  Total no. of selections of $2$ points taken at a time from $m$ points is $C_2^^m$.
\item Total no. of persons is $6 + 4 = 10$. Total no. of selections of $5$ persons out of $10$ is
  $C_5^^{10}$. Number of selections when no lady is taken is $C_5^^6$.

  Thus, no. of selections when at least one lady is present is $C_5^^10 - C_5^^6 = 252 - 6 = 246$.
\item (a) Total no. of selections of $3$ points out of $10$ points is $C_3^^{10} = 120$. Number of selections
  of $3$ points out of $4$ collinear points is $C_3^^4 = 4$.

  Thus, no. of triangles formed is $120 - 4 = 116$.

  (b) Total no. of selections of $2$ points out of $10$ points is $C_2^^{10} = 45$. No. of selection of
  points when only one line is formed is $C_2^^4 = 6$

  Therefore, no. of straight lines formed is $45 - C_2^^4 + 1 = 40$. (We take $1$ line formed from four
  collinear points)

  (c) Total no. of selections of $4$ points out of $10$ points is $C_4^^{10} = 210$. No. of selection of
  points when no quadrilateral is formed is $C_3^^4.C_1^^6 + C_4^^4.C_0^^6 = 25$.

  Thus, no. of quadrilaterals formed is $210 - 25 = 185$.
\item Zero or more oranges can be selected from $4$ oranges in $5$ ways because oranges are
  identical. Similalry, the no. of selection for apples would be $6$ and for mangoes it would be $7$.

  Thus, no. of selections when all three types of fruits are selected from is $5\times6\times7 = 210$. But
  one of these selections will contain $0$ fruits.

  Thus, required no. of selections is $209$.
\item No. of selections by which $1$ or more green dye can be chosen is $C_1^^5 + C_2^^5 + C_3^^5 + C_4^^5 +
  C_5^^5 = 2^5 - 1$. No. of selections by which $1$ or more blue dye can be chosen is $C_1^^4 + C_2^^4 +
  C_3^^4 + C_4^^4 = 2^4 - 1$. No. of selections by which $0$ or more red dye can be chosen is $C_0^^3 +
  C_1^^3 + C_2^^3 + C_3^^3 = 2^3 = 8$.

  Thus, required no. of selections is $21\times15\times8 = 3720$.
\item Factos of $216,\,000$ are $5\ 2$s, $3\ 3$s and $2\ 5$s. Zero or more $2$s can be selected in $5 + 1 =
  6$ ways. Zero or more $3$s can be selected in $3 + 1 = 4$ ways. Zero of more $5$s can be selected in $2 +
  1 = 3$ ways.

  Thus, no. of divisors is $6\times4\times3 - 1 = 71$ because one of these would contain no factor. Adding
  $1$ to the no. of divisors we have total no. of divisors as $72$.
\item A student can fail in one , two, three, four or all of five subjects. Thus, no. of ways of failing is
  $C_1^^5 + C_2^^5 + C_3^^5 + C_4^^5 + C_5^^5 = 2^5 - 1 = 31$.
\item Each person can be given $4$ things. No. of ways of giving $4$ things out of $12$ to the first person
  is $C_4^^{12}$. Then, $8$ things remain. No. of ways of giving $4$ things out of $8$ to the second person
  is $C_4^^8$. Now third person can receive $4$ things out of $4$ in $C_4^^4$ ways.

  Thus, required no. of ways is $C_4^^{12}\times C_4^^8\times C_4^^4 = \frac{12!}{(4!)^3}$.

  No. of ways in which $12$ things can be divided equally among $3$ sets is $\frac{12!}{(4!)^3.3}$.
\item There are $11$ letters in the word \quotation{Examination} in which three occur in pairs
  i.e. \quote{A}, \quote{N} and \quote{I}. The different letters are $E, X, A, M, I, N, T, O$ i.e. $8$.

  {\bf Case I:} When two pairs of identical letters are chosen.

  The two pairs can be chosen from three in $C_2^^3 = 3$ ways. These letters can be arranged among themselves in
  $\frac{4!}{2!2!} = 6$ ways. Thus, total no. of words formed is $3\times6 = 18$.

  {\bf Case II:} When one pair of identical letters is chosen and remaining two letters are different.

  The pair of identical letters can be chosen in $C_1^^3 = 3$ ways. The two different letters can be chosen
  in $C_2^^7 = 21$ ways. These letters can be arranged in $\frac{4!}{2!}$ ways.

  Thus, total no. of words formed is $3\times C_2^^7\times\frac{4!}{2!} = 756$.

  {\bf Case III:} When all four letters are different.

  No. of words that can be formed is $P_4^^8 = 1680$.

  Thus, total no. of words formed is $756 + 18 + 1680 = 2454$.
\item We need to select $4$ vertices out of $n$ of a polygon to form a quadrilateral. No. of selections of
  $4$ points is $C_4^^n$.
\item No. of ways of selecting $3$ friends out of $7$ is $C_3^^7 = 35$. Thus, no. of parties that can be
  given is $35$.

  Suppose a particular friends is mandatory in a party then $2$ other friends can be selected in $C_2^^6$
  ways. Thus, no. of parties a particular friend will attend is $C_2^^6 = 15$.
\item If $p$ things always occue then we have to select remaning $r - p$ things out of $n - p$ ways, which
  is $C_{r - p}^^{n - p}$.
\item (a) If a particular member is always added then we have to choose $5$ more from remaining $11$, which
  is $C_5^^{11}$.

  (b) If a particular member is always excluded then we have to chhose $6$ more from remaining $11$, which
  is $C_6^^{11}$.
\item (a) Total no. of ways of seating $6$ students is $P_6^^6 = 720$. Now we will put $C$ and $D$ together
  and subtract that from total no. of ways to find no. of ways of seating them when $C$ and $D$ are not
  together.

  Treating $C$ and $D$ as one student we have $5$ students which can be seated in $P_5^^5 = 120$ ways. But
  these two can be arranged among themseleves in $2!$ ways making total no. of ways $120\times2 = 240$.

  Thus, no. of ways of seating these $6$ students together when $C$ and $D$ are not together is $720 - 240 =
  480$.

  (b) If $C$ is always included then we need to select $3$ more from remaining $5$, which can be done in
  $C_3^^5 = 10$ ways.

  (c) Since $E$ is always excluded we have only $5$ students left. Thus, following previous part it can be
  done in $C_3^^4 = 4$ ways.
\item Let there be $n$ stations. To print a ticket we need a source station and a desination station. So
  different tickets which can be printed with $n$ stations is $C_2^^n$, which is $105$ in our case.

  $\therefore \frac{n!}{(n - 2)!2!} = 105 \Rightarrow n(n - 1) = 210 = 14.15 \Rightarrow n = 15$.
\item No. of ways to select $2$ points to form a straight line out of $15$ points is $C_2^^{15} = 105$.
  This will include $2$ points out of $6$ collinear points which will actually contain only $1$ straight
  line out of it. So no. of ways to choose $2$ points out of these $6$ points is $C_2^^6 = 15$. Thus, total
  no. of straight lines formed is $105 - 10 + 1 = 91$.

  No. of ways of choosing $3$ points out of $15$ is $C_3^^{15} = 455$. We have to not consider cases when
  all three points aree selected from collinear points as those won't form a triangle. No. of selections of
  $3$ points out of collinear points is $C_3^^6 = 20$.

  Thus, total no. of triangles formed is $455 - 20 = 435$.
\item No. of ways of choosing $4$ points out of $10$ is $C_4^^{10} = 210$. When $3$ or $4$ points are chosen
  from $5$ collinear pooints the quadrilateral won't be formed. When we choose $3$ points from collinear
  points we have $C_3^^5 = 10$ ways, and $1$ remaining point from $5$ non-collinear points in $5$ ways. Thus,
  total no. of such selections is $10\times5 = 50$.

  When all four points are chosen from collinear points; this can be done in $C_4^^5 = 5$ ways.

  Thus, total no. of quadrilaterals formed is $210 - 50 - 5 = 155$.
\item There is a total of $12$ points and we can choose $3$ points from these in $C_3^^{12} = 220$
  ways. However, these points must not come from points of same side.

  Thus, no. of triangles formed is $220 - C_3^^3 - C_3^^4 - C_3^^5 = 205$.
\item We need one goalkeeper in the team and two are available so goalkeeper can be chosen in $2$ ways. Rest
  of $10$ players can be chosen from remaining $12$ players in $C_{10}^^{12} = 66$ ways.

  Thus, no. of ways in which a team of $11$ out of $14$ can be formed is $2\times66 = 122$.
\item $2$ men can be chosen from $5$ men in $C_2^^5 = 10$ ways. Similarly, $2$ women from $6$ women can be
  chosen in $C_2^^6 = 15$ ways.

  Thus, total no. of ways of forming the committee is $10\times 15 = 150$.
\stopitemize