\startcomponent preface.lmtx
\setupinteraction[state=start]
\setcounter[userpage][0]

{\switchtobodyfont[sl]

\setupnotation
  [footnote]
  [numberconversion=set 2]

\starttitle
  [title=Preface]

\writetolist[chapter]{}{Preface}

\setupnotation
  [footnote]
  [numberconversion=n]

\setupheadertexts
  [\tfx Preface]
  [\tfx\pagenumber]

This is a book on algebra, which, covers basics of algebra till high school level. It covers the most essential topics to take up a
bachelor's course where knowledge of algebra is required. There is no specific purpose for writing this book. This is a book for
self study and is not recommended for courses in schools and universities. I will try to cover as much as I can and will keep
adding new material over a long period. I have no interest in writing a book in a fixed way which serves a university or college
course as I have always loved freedom. Life, freedom and honor in that order are important.

Algebra is probably one of the most fundamental subjects in Mathematics as further study of subjects like trigonometry, coordinate
geometry and rest all depend on it. That is the primary reason I have chosen it to be the first subject in mathematics to be dealt
with. It is very important to understand algebra for the readers if they want to advance further in mathematics.

\subject[title=How to Read This Book?]
Every chapter has theory. Read
that first. Make sure you understand that. Of course, you have to meet the prerequisites for the book. Then, go on and try to solve
the problems. In this book, there are no pure problems. Almost all have answers except those which are of similar kind and
repetitive in nature for the sake of practice. If you can solve the problem then all good else look at the answer and try to
understand that. Then, few days later take on the problem again. If you fail to understand the answer you can always email me with
your work and I will try to answer to the best of my ability. However, if you have a local expert seek his/her advice first. Just
that email is bad for mathematics.

Note that mathematics is not only about solving problems. If you understand the theory well, then you will be able to solve
problems easily. However, problems do help with the enforcement of theory in your mind.

I am a big fan of old MIR publisher's problem books, so I emphasize less on theory and more on problems. I hope that you find
this style much more fun as a lot of theory is boring. Mathematics is about problem solving as that is the only way to
enforce theory and find innovtive techniques for problem solving.

Some of the problems in certain chapters rely on other chapters which you should look ahead or you can skip those problems and come
back to it later. Since this books is meant for self study answers of most of the problems have been given which you can make use of.
However, do not use for just copying but rather to develop understanding.

\subject[title=Who Should Read This Book?]
Since this book is written for self study anyone with interest in algebra can read it. That does not mean that school or college
students cannot read it. You need to be selective as to what you need for your particular requirements. This is mostly high school
course with a little bit of lower classes' course thrown in with a bit of detail here and there.

\subject[title=Prerequisite]
You should have knowledge till grade 10th course. Attempt has been made to keep it simple and give as much as background to the
topic which is reasonable and required. However, not everything will be covered below grade 10.

\subject[title=Goals for Readers]
The goal of for reading this book is becoming proficient in solving simple and basic problems of algebra. Another goal would be to
be able to study other subjects which require this knowledge like trigonometry or calculus or physics or chemistry or other
subjects. If you can solve 95\% problems after 2 years of reading this book then you have achieved this goal.

All of us possess a certain level of intelligence. At average any person can read this book. But what is most important is you have
to have interest in the subject. Your interest gets multiplied with your intelligence and thus you will be more capable than you
think you can be. One more point is focus and effort. It is not something new which I am telling but I am saying it again just to
emphasize the point. Trust me if you are reading this book for just scoring a nice grade in your course then I have failed in my
purpose of explaining my ideas.

A lot of problems are given in the book for practice and you should try to solve all of these. Solutions are given to assist you
for understading. However, use them as a last resort. Slowly more and more problems will be added. There are very easy problems
which should be practiced to progress towards more difficcult problems.

Also, if you find this book useful feel free to share it with others without hesitation as it is free as in freedom.

\subject[title=Confession]
I feel like an absolute thief while writing this book for nothing given in this book is mine.
All of it belongs to others who did the original work and I have just copied shamelessly.
I have nothing new to put in the book. This book is just the result of the pain I feel
when I see young children wasting their life for they are poor. And therefore, this book
is licensed under GNU FDL. Even if I manage to create few new problems it is still based
on knowledge of other pioneers of the subject but perhaps that is how we are supposed to
progress bit-by-bit.

\subject[title=Acknowledgements]
I am in great debt of my family and free software community because both of
these groups have been integral part of my life. Family has provided direct
support while free software community has provided the freedom and freed me
from the slavery which comes as a package with commercial software. I am
especially grateful to my wife, son and parents because it is their time which
I have borrowed to put in the book. To pay my thanks from free software
community  I will take one name and that is Richard Stallman who started all
this  and is still fighting this never-ending war. When I was doing the Algebra
book then I realized how difficult it is to put Math on web in HTML format and
why Donald Knuth wrote \TeX{}. Also, \TeX{} was one of the first softwares to
be released as a free software.

Now as this book is being written using \ConTeXt{} so obviously Hans hagen
and all the people involved with it have my thanks along with Donald Knuth. I
use Emacs with Auctex and hope that someday I will use it in a much more
productive way someday.

I have used Asymptote, Metapost and Tikz for drawing all the diagrams. Both are wonderful
packages and work very nicely. Asymptote in particular is very nice for 3d-drawings and
linear equation solving. Perhaps sometime in future all diagrams will be redrawn using
Metapost.

I would like to thank my parents, wife and son for taking out their fair share
of time and the support which they have extended to me during my bad
times. After that I would like to pay my most sincere gratitude to my teachers
particularly H. N. Singh, Yogendra Yadav, Satyanand Satyarthi, Kumar Shailesh
and Prof. T. K. Basu. Now is the turn of people from software community. I must
thank the entire free software community for all the resources they have
developed to make computing better. However, few names I know and here they
go. Richard Stallman is the first, Donald Knuth, Edger Dijkstra, John von Neumann after that.
I am not a native English speaker and this book has just gone through one pair
of eyes therefore chances are high that it will have lots of errors(particularly with commas and spelling mistakes). At the
same time it may contain lots of technical errors. With time and revisiosn those errors will be removed.

\rightaligned{Dayashru(Tears of Compassion)}\\
\rightaligned{Nalanda, 2023}


\stoptitle
}
\stopcomponent
